---
title: "Homework 2"
author: Yimiao Pang
data: October 9, 2021
output: github_document
---

```{r setup, message=FALSE}
library(readxl)
library(dplyr)

library(readr)
library(tidyr)

library(tidyverse)
```



## Problem 1

the Mr. Trash Wheel sheet

```{r, message=FALSE, collapse=TRUE}
# read in
# clean_name
# omit rows that do not include dumpster-specific data
# round
tw_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() %>%
  filter(dumpster >= 1 & dumpster != "Grand Total") %>%
  mutate(sports_balls = round(sports_balls))
tw_df
```

2018 & 2019 Precipitation

```{r, message=FALSE, collapse=TRUE}
# read in
# omit rows without precipitation data
# add a variable for year

p18_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>%
  filter(Total >= 0) %>%
  mutate(Year = 2018)
p19_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>%
  filter(Total >= 0) %>%
  mutate(Year = 2019)


# combine 2018 & 2019

p_df = 
  merge(p18_df, p19_df, all = TRUE) %>%
  arrange(Year, Month) %>%
  mutate(Month = month.name[Month]) %>%
  pivot_wider(
    names_from = "Year",
    values_from = "Total"
  )
p_df
```

```{r, message=FALSE}
# filter -- trash wheel 2019
tw2019_df = filter(tw_df, year == 2019)
```

There are `r nrow(tw_df)` rows and 17 columns in trash wheel data. And there are `r nrow(p_df)` rows and 3 columns in precipitation resulting data. The variables include 2018 and 2019 and the observations are months in a year. The total precipitation in 2018 is `r sum(p18_df$Total)`, and the median number of sports balls in a dumpster in 2019 is `r median(tw2019_df$sports_balls)`.


## Problem 2

```{r 2.1, message=FALSE, collapse=TRUE}
pols_df = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(col = mon, into = c("year", "month", "day")) %>%
  mutate(month = month.name[as.numeric(month)],
         president = recode(prez_gop, '0' = "dem", '1' = "gop", '2' = "gop")) %>%
  select(-prez_dem, -prez_gop) %>%
  select(-day)
```

```{r 2.2, message=FALSE, collapse=TRUE}
#read in
snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
  separate(col = date, into = c("month", "day", "year"), sep = "/") %>%
  mutate(month = as.numeric(month),
         year = as.numeric(year))

# mutate year
snp19_df = 
  filter(snp_df, year > 15) %>%
  mutate(year = year + 1900)

snp20_df = 
  filter(snp_df, year <= 15) %>%
  mutate(year = year + 2000)

# in order -- year, month
snp_df = 
  merge(snp19_df, snp20_df, all = TRUE) %>%
  arrange(year, month) %>%
  mutate(month = month.name[month])

# leading variables -- year & month
year = select(snp_df, year)
month = select(snp_df, month)
close = select(snp_df, close)

snp_df = cbind(year, month, close)
```

```{r 2.3, message=FALSE, collapse=TRUE}
pols_df_new = 
  merge(pols_df, snp_df, by = c("year", "month"), all = TRUE)

une_df = 
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>%
  mutate(month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July", "aug" = "August", "sep" = "September", "oct" = "October", "nov" = "November", "dec" = "December"))
une_df = filter(une_df, unemployment >= 0)

pols_df_new2 = merge(pols_df_new, une_df, by = c("year", "month"), all = TRUE)
```

In the pols-month dataset, there are 9 columns and 822 rows. The time range is from January, 1947 to June, 2015.     In the snp dataset, there are 3 columns and 787 rows. The time range is from January 1950 to July 2015.     And in the unemployment dataset, there are 810 rows and 3 columns including year, month and unemployment percentage.     After joining the three datasets above, the resulting dataset has 11 columns and 823 rows. The time range is from January 1947 to July 2015. Its columns include year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem, president, close, and unemployment percentage.

## Problem 3

load and tidy

```{r, message=FALSE, collapse=TRUE}
pbn_df = 
  read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(gender = tolower(gender), 
         ethnicity = tolower(ethnicity), 
         childs_first_name = tolower(childs_first_name),
         ethnicity = recode(ethnicity, 
                             "asian and pacific islander" = "asian and paci",
                             "black non hispanic" = "black non hisp",
                             "white non hispanic" = "white non hisp")) %>%
  distinct(year_of_birth, gender, ethnicity, childs_first_name, .keep_all = TRUE)
```

Olivia

```{r, message=FALSE, collapse=TRUE}
olivia_df = 
  filter(pbn_df, childs_first_name == "olivia") %>%
  select(-gender, -childs_first_name, -count) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )
olivia_df
```

The most popular male child name

```{r, message=FALSE, collapse=TRUE}
male_df = 
  filter(pbn_df, gender == "male" & rank == 1) %>%
  select(-gender, -count, -rank) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )
male_df
```

male, white non-hispanic, 2016

```{r, message=FALSE, collapse=TRUE}
mw16_df = 
  filter(pbn_df, gender == "male" & ethnicity == "white non hisp" & year_of_birth == 2016) %>%
  arrange(rank)

ggplot(mw16_df, aes(x = rank, y = count)) + geom_point()
```

